---
title: "Overview of the blkvar package"
author: "Luke Miratrix"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library( blkvar )
library( tidyverse )
```

The `blkvar` package implements a variety of methods for estimating the average treatment impact in a blocked or multisite randomized trial. It also provides methods for estimating the amount of cross-site (or cross-block) treatment variation, defined as the variance of the individual block (or site) ATEs. It finally provides a variety of methods to implement blocked or multisite simulation studies.

## Estimating the ATE

There are several classes of methods for estimating the ATE:

* Linear regression with block dummies
* Multilevel regression
* Design based estimators
    
For a running example we generate some data
```{r}
df = gen.dat.no.cov( n.bar=50, J=30,
                         gamma.10 = 0,
                         tau.11.star = 0.2^2,
                         ICC = 0.65,
                         variable.n = TRUE )
head( df )
```

To estimate the ATE using (almost) all the various methods provided we have
```{r}
compare_methods(Yobs, Z, sid, data=df )
```

One can select which methods to use:
```{r}
compare_methods(Yobs, Z, sid, data=df, include.block = FALSE, include.MLM = FALSE,
                include.RCTYes = FALSE)
```

## Estimating cross site variation

We can estimate cross site variation using some of the multilevel modeling methods.  The `estimate.ATE.FIRC` method, for example, provides `tau.hat`, the estimate of cross site variation. To call the method, list the outcome, the treatment assignment (as a 0-1 vector, and a site identifier).
```{r}
ests <- estimate.ATE.FIRC( Yobs, Z, sid, data=df )
t( ests )
```
(We transpose with `t()` to make it a nice row instead of a list.)

This can also be done as
```{r}
ests <- estimate.ATE.FIRC( df$Yobs, df$Z, df$sid )
```


We also have
```{r}
ests.RIRC = estimate.ATE.RIRC( Yobs, Z, sid, data=df )
t( ests.RIRC )
```

These methods also have a `pool` option which uses a MLM with a single variance parameter for all individuals, rather than separate ones for the treatment and control groups (which is still pooled across block).

We can compare these estimators via
```{r}
compare_methods_variation( Yobs, Z, sid, data=df, long.results=TRUE )
```
The $p$-values are testing for these estimates being nonzero.

