---
title: "Overview of the blkvar package"
author: "Luke Miratrix"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Overview of the blkvar package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library( blkvar )
library( dplyr )
library( knitr )
```

The `blkvar` package implements a variety of methods for estimating the average treatment impact in a blocked or multisite randomized trial. It also provides methods for estimating the amount of cross-site (or cross-block) treatment variation, defined as the variance of the individual block (or site) ATEs. It finally provides a variety of methods to implement blocked or multisite simulation studies.

# Estimating the ATE

There are several classes of methods for estimating the ATE:

* Linear regression with block dummies
* Multilevel regression
* Design based estimators
    
For a running example we generate some data
```{r}
df <- gen_dat_no_cov(n.bar = 50, J = 30, gamma.10 = 0, tau.11.star = 0.2 ^ 2, ICC = 0.65, variable.n = TRUE)
head( df)
```

To estimate the ATE using (almost) all the various methods provided we have
```{r}
compare_methods(Yobs, Z, sid, data=df )
```

One can select which methods to use:
```{r}
compare_methods(Yobs, Z, sid, data=df, include.block = FALSE, include.MLM = FALSE,
                include.DB = FALSE)
```



# Adjusting for covariates

Covariate adjustment for individual level covariates is implemented for several methods (especially the linear regression based ones).

```{r}
df$X1 = df$Y0 + rnorm( nrow( df ) ) # add a fake covariate
df$X2 = df$Y1 + rnorm( nrow( df ) ) # add another fake covariate

rs = compare_methods( Yobs, Z, sid, data=df, control.formula = ~X1 + X2 )
knitr::kable( rs )
```

# Estimating cross site variation

We can estimate cross site variation using some of the multilevel modeling methods.  The `estimate_ATE_FIRC` method, for example, provides `tau.hat`, the estimate of cross site variation. To call the method, list the outcome, the treatment assignment (as a 0-1 vector, and a site identifier).
```{r}
ests <- estimate_ATE_FIRC( Yobs, Z, sid, data=df )
t( ests )
```
(We transpose with `t()` to make it a nice row instead of a list.)

This can also be done as
```{r}
ests <- estimate_ATE_FIRC( df$Yobs, df$Z, df$sid )
```


We also have
```{r}
ests.RIRC = estimate_ATE_RIRC( Yobs, Z, sid, data=df )
t( ests.RIRC )
```

These methods also have a `pool` option which uses a MLM with a single variance parameter for all individuals, rather than separate ones for the treatment and control groups (which is still pooled across block).

We can test for cross site variation using a $Q$-statistic approach:
```{r}
t( analysis_Qstatistic( Yobs, Z, sid, data=df, alpha=0.05, calc.CI=TRUE ) )
```
The confidence interval comes from inverting the test statistic for various values of cross-site variation.


We can compare all these estimators via
```{r}
compare_methods_variation( Yobs, Z, sid, data=df, long.results=TRUE )
```
The $p$-values are testing for these estimates being nonzero.

